---
title: "Introduction to dynamicAutoregCV"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to dynamicAutoregCV}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Firstly, all the necessary libraries, for the autoregression model and plots are imported.

```{r setup}
library(dynamicAutoregCV)
library(ggplot2)
```

As a use case, we work with a dataset of average daily temperature in Austin, TX. These data were extracted from the National Centers for Environmental Information (NCEI) [website](https://www.ncei.noaa.gov). 

The dataset covers the period from 2016-01-01 to 2025-08-25, with measurements in °F. As part of the data preprocessing, all NaN values were replaced with the last valid temperature measurement so that the autoregressive model is not biased by future (out-of-sample) measurements.

```{r load-data}
data("temp_austin")
str(temp_austin)
head(temp_austin)
```

